Time and space specifications for stencils

### 2016-05-24 paper.tex
preliminary model def

\newtheorem{lemma}[block]{Lemma}
\newtheorem{proposition}[block]{Proposition}
%\newtheorem{definition}[block]{Definition}

@@

% Syntax macros
\newcommand{\stenFwd}[2]{\texttt{forward} (\texttt{depth}=#1,
  \texttt{dim}=#2)}
\newcommand{\stenBwd}[2]{\texttt{backward} (\texttt{depth}=#1,
  \texttt{dim}=#2)}
\newcommand{\stenCen}[2]{\texttt{backward} (\texttt{depth}=#1,
  \texttt{dim}=#2)}

@@

We introduce a simple specification for stencil comptuations
that abstracts the fine grained detail of their access patterns.
In practise, most stencil computations have a regular shape that
can be described simply and abstractly. In the case of our
first Laplace example, our analysis tool infers the following specification:
%
\begin{SpecVerbatim}
a : centered depth=1 dims=1
b : backward depth=1\{a\} dims=t
\end{SpecVerbatim}
%
The first line explains that \texttt{a} is read from
with a symmetrical stencil pattern (``centered'') to a depth
of one in each direction in its first (and only) dimension.
The second line explains the temporal relationship between
\texttt{b} and \texttt{a}: that the previous time step for \texttt{b}
is actually provided by \texttt{a}. The dimension \texttt{t} is a
reserved dimension identifiers for temporal specifications. In
the case of the Navier-Stokes example, its inferred stencil is shown
in Figure~\ref{ref:navier-stokes-fragment}(b).

@@

\begin{figure}[t]
\begin{align*}
\def\arraystretch{1.2}
\begin{array}{rl}
\textit{spec} ::= \; & \textit{spatial} \mid \textit{temporal}   \\ \\
\textit{spatial} ::= \; & [\textit{mod} \; \{, \textit{mod} \}] \; \textit{sumS} \\
\textit{mod} ::= \; &
 \texttt{read-once} \\
& \mid \texttt{reflexive} \; \texttt{dim=}\textit{dim}{\texttt{,}} \\
& \mid \texttt{irreflexive} \; \texttt{dim=}\textit{dim}{\texttt{,}} \\
\textit{sumS} ::= \; & \textit{prodS} \;\, \texttt{+} \;\, \textit{sumS} \mid S_{\texttt{+}} \; \texttt{+}
          \; \textit{sumS} \mid \epsilon \\
\textit{prodS} ::= \; & S_{\texttt{*}}  \;\, \texttt{*} \;\, prodS
                        \mid \epsilon \\ \\
\textit{temporal} ::= \; & \texttt{dependency} \; (v \; \{ , v \})  \\ \\
\multicolumn{2}{l}{(\textit{meta rules, where \textnormal{$\odot = \{\texttt{*}, \texttt{+}, \texttt{,}\}$}})} \\
S_\odot ::= \; & \texttt{forward} \; [,\texttt{depth=}\mathbb{N}] \;
        , \texttt{dims=}\textit{dims}_\odot  \\
\mid \; & \texttt{backward} \; [,\texttt{depth=}\mathbb{N}] \; , \texttt{dims=}\textit{dims}_\odot \\
\mid \; & \texttt{centered} \; [,\texttt{depth=}\mathbb{N}] \; ,
          \texttt{dims=}\textit{dims}_\odot \\
\textit{dims}_\odot ::= & \; \mathbb{N}_{>0} \; \odot \;
                          \textit{dims}_{\odot} \; \mid \;
                          \mathbb{N}_{> 0}
\end{array}
\end{align*}
\caption{Specification syntax}
\label{fig:syntax}
\end{figure}

@@

\paragraph{Time specifications}
\dnote{The below is old so probably out of date}

\begin{ExmVerbatim}
for t = 1, itermax:
   for i = 1, (n-1):
      a[t, i] = a[t-1, i-1] - 2*a[t-1, i] + a[t-1, i+1]
\end{ExmVerbatim}

\begin{SpecVerbatim}
a(t=0) : bwd, depth=1
a(1)   : centered, depth=1
\end{SpecVerbatim}

\begin{SpecVerbatim}
a : centered depth=1 dim=1
a : backward depth=1 dim=0
\end{SpecVerbatim}

@@

\begin{ExmVerbatim}
b(x) = a(x) + r*(a(x-1) - 2*a(x) + a(x+1))
\end{ExmVerbatim}
%
For this statement, the inference provides the specification:
%
\begin{SpecVerbatim}
a: centered depth=1, dims=0
\end{SpecVerbatim}

>>>>

\newtheorem{lemma}[block]{Lemma}
\newtheorem{proposition}[block]{Proposition}
%\newtheorem{definition}[block]{Definisssstion}

@@

% Syntax macros
\newcommand{\stenFwd}[2]{\texttt{forward}, \texttt{depth}=#1,
  \texttt{dim}=#2}
\newcommand{\stenBwd}[2]{\texttt{backward}, \texttt{depth}=#1,
  \texttt{dim}=#2}
\newcommand{\stenCen}[2]{\texttt{backward}, \texttt{depth}=#1,
  \texttt{dim}=#2}

@@

We introduce a simple specification for stencil comptuations
that abstracts the fine grained detail of their access patterns.
In practise, most stencil computations have a regular shape that
can be described simply and abstractly. In the case of our
first Laplace example, our analysis tool infers the following specification:
%
\begin{SpecVerbatim}
%=  stencil centered, depth=1 dims=1 :: a
%=  stencil dependency a, mutual    :: b
\end{SpecVerbatim}
%
The first line explains that \texttt{a} is read from
with a symmetrical stencil pattern (``centered'') to a depth
of one in each direction in its first (and only) dimension.
The second line explains the temporal relationship between
\texttt{b} and \texttt{a}: that the previous time step for \texttt{b}
is actually provided by \texttt{a}. The dimension \texttt{t} is a
reserved dimension identifiers for temporal specifications. In
the case of the Navier-Stokes example, its inferred stencil is shown
in Figure~\ref{ref:navier-stokes-fragment}(b).

@@

\begin{figure}[t]
\begin{align*}
\def\arraystretch{1.2}
\setlength{\arraycolsep}{0.2em}
\newcommand{\dimTy}{\mathbb{D}}
\begin{array}{rl}
\textit{specification} ::= & [\textit{regionDec} \; \{\texttt{$\backslash$n} \;\;
                             \textit{regionDec} \}] \;
                             \textit{specDec} \\
\textit{specDec} ::= & \texttt{stencil} \; \textit{spec} \mid \epsilon
  \\
\textit{regionDec} ::= &  \texttt{region} \; \textit{rvar} \; \texttt{=} \;
                         \textit{region}\\
\textit{spec} ::= & \textit{spatial} \mid \textit{temporal}   \\[1em]
\textit{spatial} ::= & [\textit{mod} \; \{, \textit{mod} \}] \; \textit{region} \\
\textit{mod} ::= &
 \texttt{read-once} \\
\mid \; & \texttt{reflexive} \; \texttt{dims=}\dimTy \{\texttt{,} \; \dimTy\} \\
\mid \; & \texttt{irreflexive} \; \texttt{dims=}\dimTy \{\texttt{,} \;
          \dimTy\} \\[0.1em]
\textit{region} ::=
     \; & \texttt{forward} \; [,\texttt{depth=}\mathbb{N}] \; , \texttt{dim=}\dimTy  \\
\mid \; & \texttt{backward} \; [,\texttt{depth=}\mathbb{N}] \; , \texttt{dim=}\dimTy \\
\mid \; & \texttt{centered} \; [,\texttt{depth=}\mathbb{N}] \; , \texttt{dim=}\dimTy \\
\mid \; & \textit{region} \; \texttt{+} \; \textit{region} \\
\mid \; & \textit{region} \; \texttt{*} \; \textit{region} \\
\mid \; & \textit{rvar}  \\[0.5em]
\textit{temporal} ::= \; & \texttt{dependency} \; (v \; \{ , v \})
  \\[0.5em]
\dimTy ::= \; & \mathbb{N}_{>0} \\
\textit{rvar} ::= \; & [\text{\texttt{a}-\texttt{z}$\,$\texttt{A}-\texttt{Z}$\,$\texttt{0}-\texttt{9}}]+
\end{array}
\end{align*}
\caption{Specification syntax (EBNF grammar)}
\label{fig:syntax}
\end{figure}

@@

\paragraph{Temporal specifications}

@@

\begin{ExmVerbatim}
b(x) = a(x) + r*(a(x-1) - 2*a(x) + a(x+1))
\end{ExmVerbatim}
%
For this statement, the inference provides the specification:
%
\begin{SpecVerbatim}
%=  stencil centered, depth=1, dim=0 :: a
\end{SpecVerbatim}
